# Chapter 3: Logistic regression

Loading data from the url:
```{r}
library(dplyr)
url <- "https://github.com/rsund/IODS-project/raw/master/data/alc.csv"


# read the csv table
alc <- read.table(url, sep = "," , header=TRUE)
glimpse(alc)
```

This dataset is providing information from two Portugese schools regarding the students performance in Mathematics (mat) and Protugese language (por). G1 G2 and G3 are the grades in first, second period and the final grade respectively. The dataset inlcudes also information about the student background like the parents job, alcohol consuption, extra-curricular activities and others. More information is found here: https://archive.ics.uci.edu/ml/datasets/Student+Performance.

```{r}
library(ggplot2)

# initialize a plot of high_use and G3
g1 <- ggplot(alc, aes(x = high_use, y = G3))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("grade") + ggtitle("Student absences by alcohol consumption")

# initialise a plot of high_use and absences
g2 <- ggplot(alc, aes(x = high_use, y = absences))

# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ggtitle("Student absences by alcohol consumption")
# initialise a plot of high_use and health
g3 <- ggplot(alc, aes(x = high_use, y = health))

# define the plot as a boxplot and draw it
g3 + geom_boxplot() + ggtitle("Student health by alcohol consumption")
# initialise a plot of high_use and quality of family relationships
g3 <- ggplot(alc, aes(x = high_use, y = famrel))

# define the plot as a boxplot and draw it
g3 + geom_boxplot() + ggtitle("Student's quality of family relationships by alcohol consumption")


```
From the boxplots it seems that the alcohol consumption could be in a relationship with the final grade (G3), where the lower alcohol consumption could lead to better grade. Also, the high alcohol consumption could lead to more absences and students with lower alcohol consumption also seem to have a better quality of family relationships. There seems to be no relationship between the student's health and the alcohol consumption.

```{r}
#box plot for the grade by alcohol use
g1 <- ggplot(data = alc, aes(x = G3,fill = high_use))


g1 + geom_bar()

g2 <- ggplot(data = alc, aes(x = absences,fill = high_use))

# define the plot as a bar plot and draw it
g2 + geom_bar()

g3 <- ggplot(data = alc, aes(x = health,fill = high_use))

# define the plot as a bar plot and draw it
g3 + geom_bar()

g4 <- ggplot(data = alc, aes(x = famrel,fill = high_use))

# define the plot as a bar plot and draw it
g4 + geom_bar()
```

The logistic model, where the high_use is the target variable and final grade G3, absences, health and quality of family relationships (famrel) are the predictors.

```{r}
# find the model with glm()
m <- glm(high_use ~ G3 + absences+health+famrel, data = alc, family = "binomial")

# print out a summary of the model
summary(m)

# print out the coefficients of the model
coef(m)
# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI <- confint(m) %>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)

```

As predicted, the health does not seem to have any statistically significant relationship with the alcohol use. All other predictors show statistically significant relationship. The regression coefficient for the final grade shows, the log odds of the alcohol consumption is only 0,068 lower for the higher grade than for the lower.

```{r}
# define a loss function (average prediction error)
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

# compute the average number of wrong predictions in the (training) data


# K-fold cross-validation
library(boot)
cv <- cv.glm(data = alc, cost = loss_func, glmfit = m, K = nrow(alc))
cv$delta[1]
# average number of wrong predictions in the 10-fold cross validation

cv <- cv.glm(data = alc, cost = loss_func, glmfit = m, K = 10)
# average number of wrong predictions in the 10-fold cross validation
cv$delta[1]
```